---
title: "Genesis of Greek"
subtitle: "Notes"
author: "Thomas Broadwater"
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage[T1]{fontenc}
- \DeclareUnicodeCharacter{02B0}{\textsuperscript{h}}
- \DeclareUnicodeCharacter{02B7}{\textsuperscript{w}}
- \DeclareUnicodeCharacter{02C8}{\textsuperscript{\tiny|}}
- \DeclareUnicodeCharacter{02CC}{\textsubscript{\tiny|}}
- \DeclareUnicodeCharacter{2081}{\textsubscript{1}}
- \DeclareUnicodeCharacter{2082}{\textsubscript{2}}
- \DeclareUnicodeCharacter{2083}{\textsubscript{3}}
- \DeclareUnicodeCharacter{2093}{\textsubscript{x}}
- \DeclareUnicodeCharacter{2090}{\textsubscript{a}}
- \DeclareUnicodeCharacter{1E43}{\d{m}}
- \DeclareUnicodeCharacter{1E47}{\d{n}}

output: pdf_document
---

PIE voiced aspirated stops become PG voiceless aspirated stops.

\[
\begin{matrix}
  \mathrm{G^{(w)h}} \\
  \left[
    \begin{matrix}
      +~\mathrm{cons} \\
      -~\mathrm{cont} \\
      +~\mathrm{s.g.} \\
      +~\mathrm{voi}
    \end{matrix}
  \right]
\end{matrix}
\rightarrow
\begin{matrix}
  \mathrm{C^{(w)h}} \\
  \left[
    \begin{matrix}
      -~\mathrm{voi}
    \end{matrix}
  \right]
\end{matrix}
\]

```{r Aspirate Devoicing}
inventory <- c("dʰugh₂ter-", "sroˈbʰeyoh₂", "ˈh₂eǵrom", "ˈtreyes",
               "h₂ster-", "h₂enh₁mos", "ˈseptṃ")
inventory <- gsub("(d)(ʰ|ʷʰ|ʰʷ)", "t\\2", inventory, perl = T)
inventory <- gsub("(b)(ʰ|ʷʰ|ʰʷ)", "p\\2", inventory, perl = T)
inventory <- gsub("(g)(ʰ|ʷʰ|ʰʷ)", "k\\2", inventory, perl = T)
```

PIE \*s becomes PG \*h at the beginning
of a word or intervocalically

\[
\mathrm{*s}
\rightarrow
\mathrm{*h}
~/~
\left\{
  \begin{tabular}{c}
    \#\rule[-1.5pt]{1.5ex}{.5pt}V \\
    V\rule[-1.5pt]{1.5ex}{.5pt}V
  \end{tabular}
\right.
\]

PIE intervocalic \*y is elided.

```{r Yod Elision}
inventory <- gsub("(a|e|i|o|u)y(a|e|i|o|u)", "\\1\\2", inventory, perl = T)
```

\[
\mathrm{*y}
\rightarrow
\emptyset
~/~
\mathrm{V}\rule[-1.5pt]{1.5ex}{.5pt}\mathrm{V}
\]

Word initial shenanigans concerning PIE \*y
that have no solid explanation

\[
\mathrm{*s}
\rightarrow
\mathrm{*h}
~/~
\left\{
  \begin{matrix}
    \#\rule[-1.5pt]{1.5ex}{.5pt}V \\
    V\rule[-1.5pt]{1.5ex}{.5pt}V
  \end{matrix}
\right.
\]

PIE word final \*m becomes PG \*n

```{r Nasal Coronalization}
inventory <- gsub("m$", "n", inventory, perl = T)
```

\[
\begin{matrix}
  \mathrm{*m} \\
  \left[
    \begin{matrix}
      +~\mathrm{cons} \\
      -~\mathrm{son} \\
      +~\mathrm{lab} \\
      -~\mathrm{cor} \\
      +~\mathrm{nas}
    \end{matrix}
  \right]
\end{matrix}
\rightarrow
\begin{matrix}
  \mathrm{*n} \\
  \left[
    \begin{matrix}
      -~\mathrm{lab} \\
      +~\mathrm{cor}
    \end{matrix}
  \right]
\end{matrix}
~/~
\rule[-1.5pt]{1.5ex}{.5pt}\#
\]

PIE word-initial \*h₁, \*h₂, and \*h₃ become PG \*e, \*a, and \*o respectively.
```{r Initial Laryngeal Vocalization}
inventory <- gsub("^h₁", "e", inventory, perl = T)
inventory <- gsub("^h₂", "a", inventory, perl = T)
inventory <- gsub("^h₃", "o", inventory, perl = T)
```

\[
\mathrm{*h₁} \rightarrow \mathrm{*e} ~/~ \#\rule[-1.5pt]{1.5ex}{.5pt}
\]

\[
\mathrm{*h₂} \rightarrow \mathrm{*a} ~/~ \#\rule[-1.5pt]{1.5ex}{.5pt}
\]

\[
\mathrm{*h₃} \rightarrow \mathrm{*o} ~/~ \#\rule[-1.5pt]{1.5ex}{.5pt}
\]

PIE interconsonantal laryngeals show the same reflexes as the word-initial.
```{r Interconsonantal Laryngeal Vocalization}
inventory <- gsub("(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)h₁(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)",
                  "\\1e\\2", inventory, perl = T)
inventory <- gsub("(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)h₂(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)",
                  "\\1a\\2", inventory, perl = T)
inventory <- gsub("(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)h₃(t|tʰ|d|s|p|pʰ|b|k|kʰ|g|h|m|n|l|r)",
                  "\\1o\\2", inventory, perl = T)
```

\[
\mathrm{*h₁} \rightarrow \mathrm{*e} ~/~ \mathrm{C}\rule[-1.5pt]{1.5ex}{.5pt}\mathrm{C}
\]

\[
\mathrm{*h₂} \rightarrow \mathrm{*a} ~/~ \mathrm{C}\rule[-1.5pt]{1.5ex}{.5pt}\mathrm{C}
\]

\[
\mathrm{*h₃} \rightarrow \mathrm{*o} ~/~ \mathrm{C}\rule[-1.5pt]{1.5ex}{.5pt}\mathrm{C}
\]

PIE syllabic \*m and \*n may have become PG \*a by this time.

```{r Syllabic Nasal Vocalization}
inventory <- gsub('(ṃ|ṇ)', 'a', inventory, perl = T)
```

\[
\begin{matrix}
  \textrm{*ṃ~*ṇ} \\ % had to switch to make the accents work. I don't like it either.
  \left[
    \begin{matrix}
      +~\mathrm{cons} \\
      -~\mathrm{son} \\
      +~\mathrm{syl}\\
      \alpha~\mathrm{lab} \\
      \beta~\mathrm{cor} \\
      +~\mathrm{nas}
    \end{matrix}
  \right]
\end{matrix}
\rightarrow
\begin{matrix}
  \mathrm{*a} \\
  \left[
    \begin{matrix}
      -~\mathrm{con} \\
      -~\mathrm{high} \\
      +~\mathrm{low} \\
      -~\mathrm{front} \\
      -~\mathrm{back} \\
      -~\mathrm{rnd}
    \end{matrix}
  \right]
\end{matrix}
\]








